\textbf{ Accurate spectrophotometric estimates of luminosity and distance\\ for luminous red-giant stars }

### 2018-07-31 paper/sp.tex
spectral pixels and continuum

% text macros
\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\foreign}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\wise}{\project{\acronym{WISE}}}
\newcommand{\zmass}{\project{\acronym{2MASS}}}
\newcommand{\sdssv}{\project{\acronym{SDSS-V}}}

@@

We take the \apogee\ (HOGG CITE) spectral data
from \sdssiv\ (HOGG CITE) \acronym{DR14} (HOGG CITE).
Because we want to make a purely linear model, we restrict our consideration
to a small region in stellar parameter space.
We cut the \apogee\ data down to the range $0<\logg<2.2$, which isolates
stars that are more luminous than the red clump.
The \apogee\ pipeline $\logg$ values (which we use) have uncertainties but
this cut leads to a clean sample of luminous red giants, and it is a cut
that is only on spectral properties of the stars (and not photometry nor astrometry).

Because the \apogee\ target selection is based on \zmass, every \apogee\ star
comes with \zmass photometry.
That is, for each star $n$,
we have near-infrared photometry $J_n$, $H_n$, and $K_n$.
In detail we use from the catalog the entries ACE WHAT SPECIFIC CATALOG ENTRIES.

@@

Every Parent Sample object gets, in addition, a randomly assigned binary
label (A or B).
This is used for two-fold validation and jackknife.
In short, we will use the model trained on the A data to assign luminosities
and distances to the B data and \foreign{vice versa}.
The Parent Sample A has HOGG entries and Parent Sample B has HOGG entries.

For training, we need A and B Training Sets.
We define the Training Set stars to be Parent Sample stars that also
have a measured \gaia\ parallax.
The Training-Set stars, in addition to meeting all Parent-Sample cuts,
also must meet ACE WHAT ADDITIONAL CUTS to ensure that the parallax
measurements are good.
But---as we have emphasized above---we do not cut ever on parallax or
parallax error or their ratio. For this reason, most of the training set
do not have significantly measured parallaxes and HOGG percent even have
negative parallaxes!
But we will perform our training such that it will be unbiased in these
circumstances.
Indeed it would be necessarily biased if ever we did cut on parallax or
parallax signal-to-noise.

@@

>>>>

% text macros
\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\foreign}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\wise}{\project{\acronym{WISE}}}
\newcommand{\zmass}{\project{\acronym{2MASS}}}
\newcommand{\sdssiv}{\project{\acronym{SDSS-IV}}}
\newcommand{\sdssv}{\project{\acronym{SDSS-V}}}

@@

We take the \apogee\ (HOGG CITE) spectral data
from \sdssiv\ (HOGG CITE) \acronym{DR14} (HOGG CITE).
Because we want to make a purely linear model, which has very little capacity,
we restrict our consideration to a small region in stellar parameter space.
We cut the \apogee\ data down to the range $0<\logg<2.2$, which isolates
stars that are more luminous than the red clump.
The \apogee\ pipeline $\logg$ values (which we use) have uncertainties but
this cut leads to a clean sample of luminous red giants, and it is a cut
that is only on spectral properties of the stars (and not photometry nor astrometry).

From the \apogee\ data on these low-$\logg$ stars, we take the spectral pixels,
of which there are ACE HOW MANY, on a common rest-frame wavelength grid.
That is, every \apogee\ star is extracted on (or interpolated to)
the same wavelength scale.
In detail, we obtain the
the \apogee\ spectral pixels from the pipeline-generated \code{ACE whatever} files.
We then pseudo-continuum-normalize the spectra according to the procedure developed
in \project{The~Cannon} (HOGG CITE): That is, the pseudo-continuum is a spline fit to a set
of pixels chosen to be insensitive to stellar parameters.
We use as our spectral data the normalized spectral pixel values.

Because the \apogee\ target selection is based on \zmass, every \apogee\ star
also comes with \zmass photometry.
That is, for each star $n$,
we have near-infrared photometry $J_n$, $H_n$, and $K_n$.
In detail we use from the catalog the entries ACE WHAT SPECIFIC CATALOG ENTRIES.

@@

Every Parent-Sample star gets, in addition, a randomly assigned binary
label (A or B).
This is used for two-fold validation and jackknife.
In short, we will use the model trained on the A data to assign luminosities
and distances to the B data and \foreign{vice versa}.
The Parent Sample A has HOGG entries and Parent Sample B has HOGG entries.

For training, we need A and B Training Sets.
We define the Training Set stars to be Parent-Sample stars that also
have a measured \gaia\ parallax.
The Training-Set stars, in addition to meeting all Parent-Sample cuts,
also must meet ACE WHAT ADDITIONAL CUTS to ensure that the parallax
measurements are good.
But---as we have emphasized above---we do not cut ever on parallax or
parallax error or their ratio. For this reason, most of the training set
do not have significantly measured parallaxes and HOGG percent even have
negative parallaxes!
But we will perform our training such that it will be unbiased in these
circumstances.
Indeed it would be necessarily biased if ever we did cut on parallax or
parallax signal-to-noise.

@@

HOGG: How to get the data and how to use it, probabilistically.
